Paper_analyses

### 2022-02-08 Paper_Analyses_subj_obje_arousal.Rmd
bootstrap for PCA/MFA
began the fucntions to do boostrap tests to ases stability of PCA/MFA results.

we could do a regular test like the probaility of more extreme values (eigen bectors.values) to the observed ones, but not sure this is appropriate.

A common option in papers is to compute confidence intervals around the estimate, to check if the values are above a certain treshold, e,g  5% 95% for eigenvalue of 1

clustering_new_pre_PCA3.1.scale
res.mfa_bio_vs_self2_scale<- FactoMineR::MFA(clustering_new_pre_PCA3.1.scale[,2:7], 
               group = c(1, 3, 2), 
               type = c("n", "s", "s"),
               # "s", "s", "s"
               name.group = c("pp_stim","auton","subj"),
               num.group.sup = c(1),
               # ind.sup = c(1,2),
               graph = TRUE)

@@

TInPosition::
```

@@

View(wine)
FactoMineR::HMFA(clustering_new_pre_PCA3.1.scale[,2:7],
                 H = hierar, 
               # group = c(1, 3, 2), 
               type = c("n", "s", "s",  "s", "s",  "s"),
               # "s", "s", "s"
               name.group = c("pp_stim","auton","subj"),
               # num.group.sup = c(1),
                 # ncp = 
               # ind.sup = c(1,2),
               graph = TRUE)

# okay try and aggregate  response per stimuli

@@

>>>>

clustering_new_pre_PCA3.1.scale

res.mfa_bio_vs_self2_scale <- FactoMineR::MFA(clustering_new_pre_PCA3.1.scale[,2:7], 
               group = c(1, 3, 2), 
               type = c("n", "s", "s"),
               # "s", "s", "s"
               name.group = c("pp_stim","auton","subj"),
               num.group.sup = c(1),
               # ind.sup = c(1,2),
               graph = TRUE)

@@

# validation
# -	Some idea, get p values for factorminer – dimensions, and correklations
# -	Stability of firt 2 dimensions by resampling

res.mfa_bio_vs_self2_k2<- FactoMineR::MFA(clustering_new_pre_PCA3.1[,1:9], 
               group = c(4, 3, 2), 
               type = c("n", "s", "s"),
              ind.sup = c(1),
              ncp = 2,
               name.group = c("pp_stim","auton","subj"),
               num.group.sup = c(1),
               # ind.sup = c(1,2),
               graph = TRUE)

# 
# quanti

# randomly resample withouth replacement

?FactoMineR::simule()

clustering_new_pre_PCA3.1$stimDescription<- as.factor(clustering_new_pre_PCA3.1$stimDescription)

clustering_new_pre_PCA3.1$ssid_stim<- paste0(clustering_new_pre_PCA3.1$ssid,
                                             paste0(clustering_new_pre_PCA3.1$stimDescription))

clustering_new_pre_PCA3.1$ssid_stim<- as.factor(clustering_new_pre_PCA3.1$ssid_stim)

unique(clustering_new_pre_PCA3.1$ssid)

# this will simulkate 61 times (we can assum that 51 paticipants)

testsim$simul

# clustering_new_pre_PCA3.1$stimDescription

val_cutlabels<- clustering_new_pre_PCA3.1%>%
  group_by(stimDescription,val_cut_iaps) %>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  select(c("stimDescription", "val_cut_iaps"))

testsim<- FactoMineR::simule(clustering_new_pre_PCA3.1[,c(2,5:9)], 1)
# then fit pca store the
testsim_withlabels <- left_join(testsim$simul, val_cutlabels)
View(testsim_withlabels)
# then do it again a 100 times

# correlation between variable and dimension

res.mfa_bio_vs_self2_k2[["quanti.var"]][["coord"]]

res.mfa_bio_vs_self2_k2$eig[1:2,]

# https://declaredesign.org/r/fabricatr/reference/resample_data.html

resample_data(clustered_survey, N = 5, ID_labels = "clusters")

unique(clustering_new_pre_PCA3.1$ssid)
clustering_new_pre_PCA3.1[,c(1,2,5:9)]

testres<- resample_data(clustering_new_pre_PCA3.1 [,c(1,2,5:9)], N = 51, ID_labels = "ssid")

# resample_data(clustering_new_pre_PCA3.1[,c(1,2,5:9)], N = 51, 
#               ID_labels = c("ssid","stimDescription")) this can throw errior fu to unbalances

# pca
resampled_temp
# create emopty dataframe
resamp_1st<- data.frame(matrix(NA, nrow = 1, ncol = 9))
colnames(resamp_1st)<- names(resampled_temp)

# create emoty  storage rame for eigenvakues
eigen_first<- as.data.frame(eugentest)
eigen_first[,]<- NA

# eigen_first<- as.data.frame(eigen_first)

# create empty storage for correlations
cor_first<- as.data.frame(cortest)
cor_first[,]<- NA
# cor_first<- as.data.frame(cor_first)

# now store group results/partials

partials_first <- res.mfa_bio_vs_self2_k2[["partial.axes"]][["contrib"]]
partials_first[,]<- NA
partials_first

i = 1

# for i in 1:100
# set.seed(ran)

for (i in 1:1500) {
  message(sprintf("$$$$$RUNING sim %i", i))
  resampled_temp<-  resample_data(clustering_new_pre_PCA3.1 [,c(1,2,3, 5:9)], N = 51, ID_labels = "ssid")
  resampled_temp$sim_no = i

# run PCA on the resampled data
# ?list
    message(sprintf("$$$$$fitting PCA %i", i))
mfa.res_res <- FactoMineR::MFA(resampled_temp[,1:8], 
               group = c(3, 3, 2), 
               type = c("n", "s", "s"),
              # ind.sup = c(1),
              ncp = 2,
               name.group = c("pp_stim","auton","subj"),
               num.group.sup = c(1),
               # ind.sup = c(1,2),
               graph = FALSE)

# store the eigen values
  message(sprintf("$$$$$storing PCA %i", i))
eugentest<- as.data.frame(mfa.res_res$eig[1:2,])
eugentest$sim_no<- i

cortest<- as.data.frame(mfa.res_res$quanti.var$cor)
cortest$sim_no = i

partials_first<- as.data.frame(res.mfa_bio_vs_self2_k2$partial.axes$cor)
partials_first$sim_no = i
# update the initial d=frames
  message(sprintf("$$$$$upating dataframes %i", i))
resamp_1st<- (bind_rows(resamp_1st, resampled_temp))
eigen_first<- bind_rows(eigen_first, eugentest)
cor_first<- bind_rows(cor_first, cortest)
  message(sprintf("$$$$$completed iteration%i", i))
}

View(resamp_1st)
View(eigen_first)
View(cor_first)

# resamp_1st%>%
#   subset(ssid!= NA)%>%

eigen_first%>%
  mutate(component =as.factor(substr(rownames(eigen_first), 1,6)))%>%
    subset(!is.na(eigenvalue))%>%
  ggplot(aes(eigenvalue))+
  geom_histogram(alpha = .2)+
  facet_grid(~component)+
  geom_vline(xintercept = res.mfa_bio_vs_self2_k2$eig[1], linetype= "dashed")+
    geom_vline(xintercept = res.mfa_bio_vs_self2_k2$eig[2], linetype = "dotted")+
  theme_classic()

# check component 1
?quantile
quantile(eigen_first$eigenvalue,.95, na.rm = TRUE)
eigen_first%>%
  mutate(component =as.factor(substr(rownames(eigen_first), 1,6)))%>%
    subset(!is.na(eigenvalue) & component == "comp 1" )%>%
  mutate(perc_5 = quantile(eigenvalue,.05, na.rm = TRUE),
         perc_95 = quantile(eigenvalue,.95,, na.rm = TRUE))%>%
  ggplot(aes(eigenvalue))+
  geom_histogram(alpha = .2)+
    geom_density()+
  # facet_grid(~component)+
  geom_point(x = res.mfa_bio_vs_self2_k2$eig[1],y = 10,size = 2, color = "red")+
  # geom_errorbar(aes(xmin = min(perc_5), xmax=max(perc_95), ymin = 10, ymax = 12 ))
    geom_vline(aes(xintercept = perc_5), linetype = "dotted")+
  geom_vline(aes(xintercept = perc_95), linetype = "dotted")+
  theme_classic()

# dimension 2

quantile(eigen_first$eigenvalue,.95, na.rm = TRUE)
eigen_first%>%
  mutate(component =as.factor(substr(rownames(eigen_first), 1,6)))%>%
    subset(!is.na(eigenvalue) & component == "comp 2" )%>%
  mutate(perc_5 = quantile(eigenvalue,.05, na.rm = TRUE),
         perc_95 = quantile(eigenvalue,.95, na.rm = TRUE))%>%
  ggplot(aes(eigenvalue))+
  geom_histogram(alpha = .2)+
    geom_density()+
  # facet_grid(~component)+
  geom_point(x = res.mfa_bio_vs_self2_k2$eig[2],y = 10,size = 2, color = "red")+
  # geom_errorbar(aes(xmin = min(perc_5), xmax=max(perc_95), ymin = 10, ymax = 12 ))
    geom_vline(aes(xintercept = perc_5), linetype = "dotted")+
  geom_vline(aes(xintercept = perc_95), linetype = "dotted")+
  geom_errorbarh(aes(xmax = perc_95, xmin = perc_5, y = 10))
  theme_classic()

# first dimension is significant

# same thing as above with variance
quantile(eigen_first$`percentage of variance`,.95, na.rm = TRUE)
eigen_first%>%
  mutate(component =as.factor(substr(rownames(eigen_first), 1,6)))%>%
    subset(!is.na(`percentage of variance`) & component == "comp 1" )%>%
  mutate(perc_5 = quantile(`percentage of variance`,.05, na.rm = TRUE),
         perc_95 = quantile(`percentage of variance`,.95,, na.rm = TRUE))%>%
  ggplot(aes(`percentage of variance`))+
  geom_histogram(aes(y=..density..), alpha=0.5,
                 position="identity")+
  geom_density(alpha=.2) +

  # facet_grid(~component)+
      geom_errorbarh(aes(xmax = perc_95, xmin = perc_5, y = .3,ymin = .2, ymax = .4), color = "gray30")+
  geom_point(x = res.mfa_bio_vs_self2_k2$eig[1,2],y = .3,size = 3, color = "red")+

    # geom_vline(aes(xintercept = perc_5), linetype = "dotted")+
  # geom_vline(aes(xintercept = perc_95), linetype = "dotted")+
  theme_classic()

# compute a direct test
# we can check the probability of getting a vakue equal to or more extreme than our test

eigen_first%>%
  mutate(component =as.factor(substr(rownames(eigen_first), 1,6)))%>%
    subset(!is.na(`percentage of variance`) & component == "comp 1" )%>%
  mutate(extreme_eigen_log = eigenvalue>= res.mfa_bio_vs_self2_k2$eig[1])%>%
  mutate(extreme_eigen_log_prob = (sum(extreme_eigen_log))/n()) %>%
  summarise_at(c("extreme_eigen_log_prob"), mean, na.rm = TRUE)

# about 50%

# one main difference is that tests checl whether the intervak incldues a ceain value (like zero for eigen vectors or 1 for eigen valkues)
# so not sure if the test abopve is appropriate withouth correcting for reversals

```{r}
eigen_first%>%
  mutate(component =as.factor(substr(rownames(eigen_first), 1,6)))%>%
    subset(!is.na(eigenvalue) & component == "comp 1" )%>%
  ggplot(aes(`percentage of variance` ))+
  geom_histogram(alpha = .2)+
  # facet_grid(~component)+
  geom_vline(xintercept = res.mfa_bio_vs_self2_k2$eig[1,2], linetype= "dashed")+
    # geom_vline(xintercept = res.mfa_bio_vs_self2_k2$eig[2], linetype = "dotted")+
  theme_classic()

eigen_first%>%
  mutate(component =as.factor(substr(rownames(eigen_first), 1,6)))%>%
    subset(!is.na(eigenvalue) & component == "comp 1" )%>%
  ggplot(aes(`percentage of variance` ))+
  geom_histogram(alpha = .2)+
  # facet_grid(~component)+
  geom_vline(xintercept = res.mfa_bio_vs_self2_k2$eig[1,2], linetype= "dashed")+
    # geom_vline(xintercept = res.mfa_bio_vs_self2_k2$eig[2], linetype = "dotted")+
  theme_classic()

# let's gp to variables

cor_first %>%
  mutate(variable = as.factor(substr(rownames(cor_first), 1,6)))%>%
    subset(!is.na(Dim.1)) %>%
  subset(variable == "arousa"|variable == "valenc")%>%
  # select(c("arousal", "valence"))%>%
  ggplot(aes(Dim.1))+
  geom_histogram(alpha = .2)+
  facet_grid(~variable)+
  geom_vline(xintercept = .75)+
    # geom_vline(xintercept = res.mfa_bio_vs_self2_k2$eig[2], linetype = "dotted")+
  theme_classic()

cor_first %>%
  mutate(variable = as.factor(substr(rownames(cor_first), 1,6)))%>%
    subset(!is.na(Dim.1)) %>%
  subset(variable != "arousa"|variable != "valenc")%>%
  # select(c("arousal", "valence"))%>%
  ggplot(aes(Dim.2))+
  geom_histogram(alpha = .2)+
  facet_grid(~variable)+
  geom_vline(xintercept = .58)+
    # geom_vline(xintercept = res.mfa_bio_vs_self2_k2$eig[2], linetype = "dotted")+
  theme_classic()

 res.mfa_bio_vs_self2_k2$quanti.var$cor

unique(as.character(testres$ssid))

mean(testres$pup_basCor)
mean(clustering_new_pre_PCA3.1$pup_basCor)

testres%>%
  ggplot(aes(valence, arousal))+
  geom_point()+
  ggpubr::stat_cor()

clustering_new_pre_PCA3.1%>%
  ggplot(aes(valence, arousal))+
  geom_point()+
  ggpubr::stat_cor()

# okay this fucntion seems to work
```

@@

View(wine)

@@

?FactoMineR::MFA
View(res.mfa_val_cat_scal)

# force k = 2

res.mfa_val_cat_scal_k2<- FactoMineR::MFA(clustering_new_pre_PCA3_val_vat_scale_then_agg1,
               group = c(2, 3, 2), 
               type = c("n", "s", "s"),
               name.group = c("ss_id_stim","auton","subj"),
               num.group.sup = c(1),
               ncp = 2,
               # ind.sup = c(1,2),
               graph = TRUE)

res.mfa_val_cat_scal_k2$eig
